Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-07 test_extension.Rmd
edits to text

```{r rp_aggr_plot, echo = FALSE, fig.cap = "This plot examines the pattern of missing data in the the original racial polarization dataset from Trounstine (2016). The plot on the left shows the percentage of missing data for each variable with missing data and reveals that the variables that are missing by far the most data are the largest vote split along racial lines (biggestsplit) and the variable indicating whether or not the election had candidates of more than one race (biracial). The plot on the right shows the pattern of missingness among the variables. We can see that approximately 42\\% of observations are complete (have no missing values). There seems to be a correspondence between missing a value for biggest split and missing biracial. There also seem to be about 2\\% of values for which most of the variables are missing. However, most observations are not missing more than 2-3 values.", fig.width= 6, fig.height = 5, warning = F, fig.pos="H"}

# select only the missing values

rp_aggr <- rp_impute %>%
   select("biggestsplit", "biracial", "whiteideology_fill2", "medincinterp",
            "diversityinterp",
             "pctasianpopinterp", "pctblkpopinterp", "pctlatinopopinterp",
             "pctrentersinterp", "pctcollegegradinterp", "nonpartisan",
               "south", "midwest", "west") %>%
  rename("Racial Divide" = biggestsplit, "Biracial Contest" = biracial, 
         "White Ideology" = whiteideology_fill2, "Median HH Income" = medincinterp,
         "Diversity" = diversityinterp, "% Asian" = pctasianpopinterp, 
         "% Black" = pctblkpopinterp, "% Latino" = pctlatinopopinterp,
         "% Renters" = pctrentersinterp, "% College Grads" = pctcollegegradinterp,
         "Nonpartisan Elect." = nonpartisan, "South" = south, "Midwest" = midwest,
         "West" = "west")

# create an aggr plot for the missing values
par(font = 1)
aggr(rp_aggr, bars = TRUE,  col=c('navyblue','red'), numbers=TRUE, cex.axis=.5, gap=1, ylab=c("Proporiton of Missing Values","Combinations"))
title(main = "Patterns of Missingness in the Financial Segregation Data",
      sub = "", line = 3, oma = c(1,3,5,1), cex.axis = .5, font.main = 1, font.axis = 1, cex.main = 1, cex.lab = .75)

@@

```{r rp_densityplot, echo = FALSE}

densityplot(imp_1_rp, main = "Density Plots of Imputed and Original Data by Variable\nin the Racial Polarization Dataset",
      par.settings = my.settings,
      par.strip.text = list(cex = .7))

@@

Since the results of the diagnostic checks were promising, I next proceeded to re-estimate the original model using the multiply imputed datasets, pooling the results to produce final pooled regression coefficients and parameters. The results for the analyses using the racial polarization dataset are presented in tables X1-X3.

```{r rp_imputations, include = FALSE}

# Now that we have imputed the missing data, we can re-run the analyses from the
# original paper to see how and if this changes the analysis. We can use the
# with() function from the mice package to fit each of the imputed datasets to
# the model and then pool the results afterwards.

# regression 1: fit multiple imputed datasets

fit_imp_1_rp <- with(imp_1_rp, lmer(formula = biggestsplit ~ H_citytract_multi_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_1_rp <- pool(fit_imp_1_rp)
imp_1_rp_sum <- summary(pool_imp_1_rp)

## regression 2 Table 1: fit multiple imputed datasets

#formula = biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2)

fit_imp_2_rp <- with(imp_1_rp, lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_2_rp <- pool(fit_imp_2_rp)
imp_2_rp_sum <- summary(pool_imp_2_rp)

## regression 3 Table 1 with new data

#formula = biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2)

fit_imp_3_rp <- with(imp_1_rp, lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_3_rp <- pool(fit_imp_3_rp)
imp_3_rp_sum <- summary(pool_imp_3_rp)

@@

```{r results_printing_rp_impute_regressions_1, results = "asis", echo=FALSE}

imp_1_rp_sum_table <- as.data.frame(imp_1_rp_sum) %>% 
  select(term, estimate, std.error, p.value) %>%
  head(13) %>%
  mutate()

imp_1_rp_sum_table_vars <- tibble(
  Variable = c("Intercept", "Multigroup Seg. Index", "Diversity", "% Asian", "% Black", "% Latino",
               "Median Income (1000s)", "% Renters", "% College Grads", "Biracial Elec.", "Nonpartisan Elec.", "Primary Elec.", "Pop (logged)"),
  Estimate = round(imp_1_rp_sum_table$estimate, 3),
  `Std. Error` = round(imp_1_rp_sum_table$std.error,3),
  `P-Value` = round(imp_1_rp_sum_table$p.value, 3)
  )

kable(imp_1_rp_sum_table_vars, "latex", caption = "\\textbf{Racial Polarization in Segregated Cities: Racial Divide with Multigroup Segregation Index}", booktabs = T) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's first model for the effect of segregation on political polarization",
  "(racial vote divide) using the multi-group Theil's H segregation Index using the new, multiply imputed",
  "datasets. The sign of the coefficient on the main variable of interest, the segregation index, is the same, but",
  "the result has become statisically insignificant with the 95% confidence interval containing zero. Thus",
  "while these results still suggest that segregation may be associated with increased political polarization,",
  "they confer a lesser degree of certainty than Trounstine's original analysis. The coefficients for % Black",
  "population, % Latino population, and median household income have also switched signs. The coefficient",
  "on the indicator for a primary election, has also changed from being significant in the original analysis",
  "to insignificant here. In general, the standard errors on the coefficients have decreased slightly."),
           general_title = "Table 1:")%>%
  kable_styling(full_width = TRUE)

@@

```{r results_printing_rp_impute_regressions_2, results = "asis", echo=FALSE}

imp_2_rp_sum_table <- as.data.frame(imp_2_rp_sum) %>% 
  select(term, estimate, std.error, p.value) %>%
  head(13)

imp_2_rp_sum_table_vars <- tibble(
  Variable = c("Intercept", "White/Nonwhite Seg. Index", "Diversity", "% Asian", "% Black", "% Latino",
               "Median Income (1000s)", "% Renters", "% College Grads", "Biracial Elec.", "Nonpartisan Elec.", "Primary Elec.", "Pop (logged)"),
  Estimate = round(imp_2_rp_sum_table$estimate, 3),
  `Std. Error` = round(imp_2_rp_sum_table$std.error,3),
  `P-Value` = round(imp_2_rp_sum_table$p.value, 3)
  )

kable(imp_2_rp_sum_table_vars, "latex", caption = "\\textbf{Racial Polarization in Segregated Cities: Racial Divide with Two-Group Segregation Index}", booktabs = T) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's second model for the effect of segregation on political polarization",
  "(racial vote divide) using the two-group Theil's H segregation Index using the new, multiply imputed datasets.", 
  "Compared to Trounstine's original results, the main coefficient on segregation index has decreased in magnitude",
  "although, it remains statistically significant. The other coefficients except for the biracial indicator are",
  "insignificant, and the coefficients on percent Black, percent Latino, and Median Income have switched signs.",
  "Standard errors have also reduced slightly in these specifications compared to the original analysis in",
  "Trounstine (2016)."),
           general_title = "Table 2:") %>%
  kable_styling(full_width = TRUE)

@@

```{r results_printing_rp_impute_regressions_3, results = "asis", echo=FALSE}

imp_3_rp_sum_table <- as.data.frame(imp_3_rp_sum) %>% 
  select(term, estimate, std.error, p.value) %>%
  head(14)

imp_3_rp_sum_table_vars <- tibble(
  Variable = c("Intercept", "White/Nonwhite Seg. Index", "Diversity", "% Asian", "% Black", "% Latino",
               "Median Income (1000s)", "% Renters", "% College Grads", "Biracial Elec.", "Nonpartisan Elec.", "Primary Elec.", "Pop (logged)", "White Ideology"),
  Estimate = round(imp_3_rp_sum_table$estimate, 3),
  `Std. Error` = round(imp_3_rp_sum_table$std.error,3),
  `P-Value` = round(imp_3_rp_sum_table$p.value, 3)
  )

kable(imp_3_rp_sum_table_vars, "latex", caption = "\\textbf{Racial Polarization in Segregated Cities: Racial Divide with Ideology Control}", booktabs = T) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's third model for the effect of segregation on political polarization",
  "(racial vote divide) using the two-group Theil's H segregation Index and including controls for ideology",
  "using the new, multiply imputed datasets. Compared to Trounstine's original results, the main coefficient on",
  "segregation index has decreased in magnitude although, it remains statistically significant. The other",
  "coefficients except for the biracial indicator are insignificant, and the coefficients on percent Asian and",
  "Black have switched signs. Standard errors have also reduced slightly in these specifications compared to",
  "the original analysis in Trounstine (2016)."),
           general_title = "Table 3:") %>%
  kable_styling(full_width = TRUE)

@@

As a result of the imputation process, I was able to include an additional XXXXXX observations in the analysis.

@@

@@

Because these results differed to such a degree from @trounstine_segregation_2016's original findings, I performed some checks on the data included in each analysis. I discovered that @trounstine_segregation_2016 excluded most cities below a certain population size because of the nature of the paper's listwise deletion. It may be that small cities are more likely to be missing data, and thus they were more likely to be dropped in @trounstine_segregation_2016's analysis. Referring back to @baqir_public_1999, I decided to divide the orginal sample based upon the populaiton cutoff for cities used in that paper, which determined that the populaiton must be above 25,000 people for an area to be considered a city. Splitting the financial segregation dataset into two datasets, one above and one below the 25,000 population cutoff, I then imputed missing values and used the multiply imputed datasets in Trounstine's original models.

Figure XXX displays the populaiton distributions used in each of these analyses. The figure makes it clear that the observations included in @trounstine_segregation_2016's original analysis were not representative of the original datasample with the missing values, as most of the cities with lower populations were excluded as a result of the list-wise data deletion. While the dataset generated from the data imputation on the whole sample also has a different distribution for city populaiton, the range of the data seems to be much more representative of the overall data range in comparison to the @trounstine_segregation_2016 sample. For comparison, I also include the distributions of the multiply imputed datasets above and below the 25,000 populaiton cutoff. For cities above the cutoff, the city population distribution is similar to the distribution in @trounstine_segregation_2016.

@@

@@

The results of the models run with the multiply imputed datasets above and below the cutoff are presented in the following two tables.

imp_over_baqir_1_sum_table_vars <- tibble(
  Variable = c("White/Nonwhite Seg. Index", "Diversity", "% Black", "% Asian", "% Latino", "Median Income (1000s)", "% Local Gov. Worker", "% Renters", "% 65+", "% College Grads", "Pop (logged)"),
  Estimate = round(imp_over_baqir_1_sum_table$estimate, 3),
  `Std. Error` = round(imp_over_baqir_1_sum_table$std.error,3),
  `P-Value` = round(imp_over_baqir_1_sum_table$p.value, 3)
)

As a summary of all of the model results for the two main coefficients of interest, segregation index and diversity, a coefficient plot is presented in Figure XXX, with the models ordered by increasing median log population size of the sample on the x-axis. Looking first at the coefficients for diversity, the two models with low median city populaiton size have significant, positive coefficients, while the models with the larger median city populations have insignificant coefficients. This suggestst that diversity is an important estimator of public goods spending for small, but not necessarily for large, cities. Next, looking at the plot for the segregation index, we see the oppposite effect. For the models where the median city population sizes were small, the coefficient on segregaiton is insignificant. However, for the two models with larger median city populations, the coefficients were significant and negative. This suggests that for small cities, segregation is not a very good estimator of public goods spending, but when cities are larger, increases in a city's segregation index are associated with decreases in public goods spending.

@@

The results of the data imputation exercise differed between the two datasets. For the analyses using the smaller, racial polarization dataset, the results from the imputed data resulted in a smaller magnitude of the coefficient on the main independent variable, and also made it statistically insignificant in the main model specification. This suggests that the segregation index may not have as strong of a positive association with political polarization as @trounstine_segregation_2016 suggests. Nevertheless, the imputation exercise generally supports the main ideas in this part of Trounstine's analysis.

On the other hand, for the analyses using the much larger financial segregation dataset, the results of the models using the multiply imputed data sharply contrasted the results of @trounstine_segregation_2016, suggesting that diversity, and not segregation, was the best explanatory variable for public goods spending across cities. While these results originally seemed quite contradictory, I was able reconcile this finding by discovering that the results of @trounstine_segregation_2016 seem to be only really representative of larger cities. By dropping all observations in the original data with missing values for key variables, it seems that @trounstine_segregation_2016 inadvertently restricted their sample to almost entirely larger cities (above 25,000 residents) whereas the full sample consisted of cities that had smaller populaitons. To test the hypothesis that diversity is more important for predicting public goods spending for small cities while segregation is more important for predicting public goods spending for larger cities, I divided the original dataset into two subsets based on the city classificaiton criteria used in @baqir_public_1999 that cities must have a population of at least 25,000 people. I then multiply imputed the missing values for each data subset and re-ran the models. I found that the results presented initial evidence in support of my hypothesis. for the cities below the 25,000 person populaiton cutoff, the coefficient for diversity was positive and significant while the coefficient for the segregation index was insignificant. For cities above the 25,000 person cutoff, the results were very similar to the results found in @trounstine_segregation_2016: the coefficient for diversity was insignificant, but the coefficient for the segregation index was significant and negative.

The replication exercise and the robustness check of @trounstine_segregation_2016's model using multiply imputed data provide additional evidence that residential racial segregation is associated with political polarization. Furthermore, I find evidence that supports the finding of @trounstine_segregation_2016 that increases in segregation are associated with decreases in public goods spending; however, I find that this is only the case for larger cities. In contrast to the argument made by @trounstine_segregation_2016, I find that, for small cities, diversity, not segregation, seems to be the most important explanatory variable for public goods spending. One potential explanation for this result could be that when cities become too small, the sort of segregation by neighborhood described in @trounstine_segregation_2016 as a diverse city broken down into homogenous, segregated neighborhoods becomes hard to maintain. When populaiton sizes are small, segregation may become difficult and inter-racial interactions may become more common, leading diversity to be more indicative of a degree of integration. This provides a possible explanation as to why the coefficient on diversity was positive for very small cities, as these "cities" may function more closely as neighborhoods (many of them with populaitons ranging from X-X thousand); as discussed in the literature review, diversity at the neighborhood level is associated with lower levels of political polarization (Oliver 2010) and thus we would expect these areas to have higher levels of public goods spending according to the ideas of @trounstine_segregation_2016. As cities become larger and larger, however, comprehensive segregation becomes more attainable. Thus, as cities become larger, segregation, not diversity, becomes a better indicator of inter-racial harmony and cooperation and thus the best explanatory variable for public goods spending.

It is possible that other mechanisms may be more important in explaining why segregation might be associated with decreased spending on public goods. For example, it could be that there is an important interaction effect between income and segregation, as others have previously suggested, or there could be some other intervening mechanism or a confounding variable that is leading to the negative association between segregation and lower spending on public goods. @trounstine_segregation_2016 has identified an important relationship between segregation and the provision of public goods. However, my analysis suggests that further research is needed to assess the potential mechanisms by which residential segregation by race or even a confounding variable may result in dimished spending. If this mechanism can be better understood, than it will be possible to more effectively understand and address the underprovision of public goods, which may inhibit social mobility and unfairly disadvantage certain groups.

# Appendix

```{r  include = FALSE}

## regression 1 Table 1

# perform a multi-level mixed-effects linear regression of biggest split on the multi-group H index with controls for Herfindahl Diversity Index, % Asian interpolated, % Black interpolated, % Latino interpolated, median household income interpolated, % renters interpolated, percent college grads interpolated, a dummy indicating if there were biracial candidates in the election, a dummy  indicating if the election was nonpartisan, an indicator if the election was a primary, and the log population. The regression also includes fixed effects for year and region and random effects for cities (geo_id2). Diversity is  included because of the claim that more diverse populations will have more heterogeneous political preferences. The other fixed effect control variables are included because they are known to be potentially correlated with segregation. winner == 1 requires that there was a winner in the election.

# regression specification. Apply REML = FALSE. The default for lmer() in R is
# to optimize the restricted (residual) maximum likelihood (REML) whereas in
# xtmixed in Stata the default is to optimize the log likelihood (which you can
# do in lmer() in R by setting REML = FALSE), so this will match the STATA code

m1 <- lmer(biggestsplit ~ H_citytract_multi_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

## regression 2 Table 1

# Perform the same regression as for Column 1 above but this time regress on the two-group calculation of Theil's H interpolated. It is unclear where the "diversity" variable is coming from, as the only variable related to diversity in the specificed dataset is diversityinterp. This is a problem that will need to be resolved. Essentially, this regression just tests an alternate specification to the first in which the segregation index in defined in terms of two-groups rather than multiple.

m2 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

# perform the same regression as for column 2, but now add in a measure of
# average white political ideology. The point of this is to control for ideology
# so that the author can separate any effects of segregation from any potential
# effects of white political ideology. This is important because it is possible
# that segregation is just a proxy for white conservatism and thus failure to
# include this variable could lead to ommitted variable bias.

m3 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

```{r results="asis", echo = FALSE}

# create a stargazer table of the results

t1 <- stargazer(m1, m2, m3, omit = c("year", "south", "midwest", "west"),
                header = FALSE,
                style = "apsr",
                title = "\\textbf{Racial Polarization in Segregated Cities}",
                single.row = FALSE,
                no.space = F,
                covariate.labels = c("Multigroup H Index", "White/Nonwhite H Index", "Diversity", "Percent Asian", "Percent Black", "Percent Latino", "Medain HH Income (1000s)", "Percent Renters", "Percent College Degree", "Biracial Contest", "Nonpartisan Election", "Primary Election", "Population (logged)", "White Ideology", "Constant"),
                dep.var.labels = c("\\textbf{\\underline{Racial Divide with:    }}"),
                column.labels = c("Multigroup H Index", "White/Nonwhite H Index", "Ideology Controls"),
                notes = c("Table A1: This table displays the results of the replication effort of the analyses conducted in Trounstine (2016)", 
                          "using the racial polarization dataset. The replication code can be found on my personal Github repository, which is",
                          " linked in the introduction. The original table is displayed for comparison as Figure 3."),
                table.placement = "H",
                notes.align = "l",
                font.size = "small")

![This is the original table displaying the results of the models for the effects of residential segregation on political polarization from page 713 of @trounstine_segregation_2016. My replication of these results can be found in Table A1 in this paper.](table_1.png)

```{r felm, include = FALSE}

# Main Analysis 2

# The following is a fixed effects model regression with clustering around
# geo_id2. It mandates that total census tracts in the city is greater than 1,
# which we want because our measure of evenness of racial spread is constant for
# cities with only one tract by our definition (which requires comparing tracts
# within cities). It also mandates that Direct General Expenditures (DGE) per
# capita, CPI adjusted is not equal to 0, which would be problematic and
# indicate a potential data error. This regression regresses the DGE per capita
# on the the two-group calculation of  Theil's H interpolated. Controls are
# included for diversity and the % populations of Blacks, Asians, and Latinos to
# help identify whether it is actually diversity that drives down spending.These
# demographic controls also help to account for the fact that whites and
# minorities tend to perfer different levels of government spending in general.
# This will help isolate the effects of segregation v. other forms of diversity
# on public spending. Other controls for Median household income, CPI adjusted,
# percent over 65, percent college grads, percent local government workers, log
# population help control for additional factors related to segregation and
# expenditure. Controls for percent renters interpolated and Median household
# income also help to control for the general wealth of cities, which could be a
# potential driver of public spending. Fixed effects for cities are also
# included  so that the author can examine the effect of segregation in the same
# city over  time, which also helps control for many other factors not otherwise
# taken into account (such as city age)

# apply the conditions in Stata regression

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), dgepercap_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

# felm(y~x | factor(fe1) + factor(fe2) | 0 | clustervar)

felm1 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# This regression is the same as above except instead of using diversity, a new
# control is added for the five year changes in racial group shares. The goal
# here is to see whether or not changes in diversity are a driving factor
# (rather than absolute levels of diversity).

felm2 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + chng5pctblk + chng5pctlatino + chng5pctasian+ medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 3

# As in column 3 of table 1, this regression simply adds in a control for  mean  ideology of city residents from General Social Survey (GSS) to control for the  possibility that ideology that segregated cities are more ideologically  conservative and that the conservative nature of segregated cities is what actually drives the effect.

felm3 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop + ideology_fill | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

```{r main_fp_tables, results = "asis", echo = FALSE}

# create main table for table 3

table2 <- stargazer(felm1, felm2, felm3, omit = c("year", "south", "midwest", "west"), 
                    header = FALSE,
                    style = "apsr",
                    title = "\\textbf{Effect of Segregation on Overall per Capita City Expenditures}",
                    covariate.labels = c("White/Nonwhite H Index", "Diversity", "Percent Black", "Percent Asian", "Percent Latino", "Change Percent Black", "Change Percent Latino", "Change Percent Asian", "Medain HH Income (1000s)", "Percent Government Worker", "Percent Renters", "Percent 65+", "Percent College Degree", "Population (logged)", "White Ideology", "Constant"),
                    dep.var.labels = c("\\textbf{\\underline{Direct General Expenditure Per Capita}}"),
                    column.labels = c("Main Specification", "Changing Demographics", "Ideology Controls"),
                    notes = c("Table A2: This table displays the results of the replication effort of the analyses conducted in Trounstine (2016)", "
                              assessing the effect of segregation in direct per capita public expenditures using the financial segregation",
                              "dataset. The replication code can be found on my personal Github repository, which islinked in the introduction."),
                    table.placement = "H",
                    notes.align = "l",
                    font.size = "small")

```{r, include = FALSE}

# Main Analysis 3

## regression 1 Table 3

# Column 1

# The following is a fixed effects model regression with clustering once again around geo_id2. Again, the if statements at the end mandate that total census tracts in the city is greater than 1, which we want because our measure of evenness of racial spread is constant for cities with only one tract by our definition (which requires comparing tracts within cities). They also mandate that the highways per capita, CPI adjsuted with no cap extend (lagged 5 years) is not equal to 0, which would be problematic and indicate a potential data error. The dependent variable here is highways per capita, CPI adjsuted with no cap extend (lagged 5 years) and it is regressed on the two group calculation of Theil's H, interpolated. Controls are added for diversity, percent population of Blacks, Asians, and Latinos, and median household CPI adjusted income as well as for % local government worker hundreds, percent rentership, percent over 65, percent college graduates, and the log of the population. This essentially examines the effect of segregation on the specific public good of highways. As in the previous table's regressions, fixed effects for cities are also included so that the author can examine the effect of segregation in the same city over time, which also helps control for many other factors not otherwise taken into account (such as city age) *

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), highwayspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_1 <- felm(highwayspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

## regression 2 Table 3

# Column 2

# This is the same regression as column 1, but this time the dependent variable
# is police per capita, CPI adjusted with no cap extend. This essentially
# examines the effect of segregation on the specific public good of police per
# capita.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), policepercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_2 <- felm( policepercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 3
fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), parkspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

# This is again the same regression, but this time the dependent variable is
# parks per capita, CPI adjusted with no cap extend. This essentially examines
# the effect of segregation on the specific public good of parks.

felm2_3 <- felm(parkspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 4

# This is again the same regression, but this time the dependent variable is
# sewers per capita, CPI adjusted with no cap extend. This essentially examines
# the effect of segregation on the specific public good of sewers.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), sewerspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_4 <- felm(sewerspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 5

# This is again the same regression, but this time the dependent variable is
# Welfare, Health, Housing per capita, CPI adjusted, no cap expend. This
# essentially examines the effect of segregation on the specific public good of
# welfare and housing.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), welfhoushealthNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_5 <- felm(welfhoushealthNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 6

# This is again the same regression, but this time the dependent variable is own
# source of general revenue per capita, CPI adjusted. This essentially examines
# the effect of segregation on the specific public good of own source revenue.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), genrevownpercap_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_6 <- felm(genrevownpercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

```{r iv_regs, include = FALSE}

## regression 1 Table 5

# Top Row

# Column 1

# This is an instrumental variable regression with DGE per capita, adjusted by the CPI, as the dependent variable and total number of waterways as the instrumental variable for segregation. The instrument includes the log of the population in the first stage because the number of waterways is positively correlated to population and population is positively correlated with segregation. The control variables are the same as those in the regressions for Table 2 except for 2 changes: since the number of waterways does not change, there is no need for city fixed effects. Instead, we include fixed effects for the region and year. A lagged version of the dependent variable (dgepercap_cpilag) is added to the regressions to account for the high correlation between  observations for the same city over time and because changes in local budgets typically are incremental (NOTE: why exactly do we do this/why does it help/why not do it in the non-intstrumental regressions?).

fin_seg_1 <- fin_seg %>%
  filter(dgepercap_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

glimpse(fin_seg_1$medincinterp)

# good explanation: http://eclr.humanities.manchester.ac.uk/index.php/IV_in_R

# iv = ivreg(second stage regression | instrument1 + instrument2 )

ivreg1 <- ivreg(dgepercap_cpi ~ H_citytract_NHW_i + dgepercap_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + dgepercap_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_1)

# Column 2

# this is the same regression as for column 1 but now we use the dependent variable highwayspercapNC_cpi to assess the effect of the instrumental  variable on highways per capita *

fin_seg_2 <- fin_seg %>%
  filter(highwayspercapNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

>>>>

@@

@@

Given the promising results of the diagnostic checks, I next proceeded to re-estimate the original model using the multiply imputed datasets, pooling the results to produce final pooled regression coefficients and parameters. The results for the analyses using the racial polarization dataset are presented in tables X1-X3.

@@

@@

@@

@@

@@

The results of the first regression indicate that Trounstine's results are robust to the inclusion of the imputed data, which led to the inclusion of an additional 59,377 observations in the analysis. The coefficients, standard errors, and significance levels are essentially unchanged. The same is largely true for the second and third specifications, but the magnitude of the coefficient on the segregation decreases slightly.

@@

@@

imp_over_baqir_1_sum_table_vars <- tibble(
  Variable = c("White/Nonwhite Seg. Index", "Diversity", "% Black", "% Asian", "% Latino", "Median Income (1000s)", "% Local Gov. Worker", "% Renters", "% 65+", "% College Grads"),
  Estimate = round(imp_over_baqir_1_sum_table$estimate, 3),
  `Std. Error` = round(imp_over_baqir_1_sum_table$std.error,3),
  `P-Value` = round(imp_over_baqir_1_sum_table$p.value, 3)
)

@@

@@

The results of the data imputation exercise differed between the two datasets. For the analyses using the smaller, racial polarization dataset, the results from the imputed data resulted in a smaller magnitude of the coefficient on the main independent variable, and also made it statistically insignificant in the main model specification. This suggests that the segregation index may not have as strong of a positive association with political polarization as @trounstine_segregation_2016 suggests. However, for the analyses using the much larger financial segregation dataset, the results of the models using the multiply imputed data largely mirrored the results of @trounstine_segregation_2016, and the coefficients and significance levels were essentially the same between the two, with the standard errors on the model estimates being generally slightly lower for the analyses using the imputed data. This provides strong support for @trounstine_segregation_2016's finding that residential racial segregation is associated with diminshed spending on public goods.

The replication exercise and robustness check using multiply imputed data here provide strong evidence that residential segregation along racial lines does in fact correlate negatively with spending on a variety of public goods. What my analysis calls more into question, is by what mechanism this happens. Using the multiply imputed data, I still found some evidence that segregation may be associated with increased political polarization which @trounstine_segregation_2016 suggests may lead to diminished public goods spending; however, the effect size is much smaller than that found in @trounstine_segregation_2016, and some of the results have become insignificant. It is possible that other mechanisms may be more important in explaining why segregation might be associated with decreased spending on public goods. For example, it could be that there is an important interaction effect between income and segregation, as others have previously suggested, or there could be some other intervening mechanism or a confounding variable that is leading to the negative association between segregation and lower spending on public goods. @trounstine_segregation_2016 has identified an important relationship between segregation and the provision of public goods. However, my analysis suggests that further research is needed to assess the potential mechanisms by which residential segregation by race or even a confounding variable may result in dimished spending. If this mechanism can be better understood, than it will be possible to more effectively understand and address the underprovision of public goods, which may inhibit social mobility and unfairly disadvantage certain groups.
